Page not found

### 2023-06-20 manuscript/appendices/B_mean_patterns.qmd
so... many.... figures !

::: {.content-hidden}
{{< include ../macros.tex >}}
\def\A{\mathcal{A}}
\def\B{\mathcal{B}}
:::

@@

* @fig-kernels-samplingR2 is the method described in this appendix, based on sampling $N$ neurons in the PDF space $\R^2$, applying the mapping on the obtained positions $\vec{z_i}$, the applying the binning (downsampling). We see, through the empty rows in the connectivity, that due to finite-$N$ effect, locations where $\rho$ is small are not represented, and, as a consequence, no neurons are present in the corresponding 1D bin. When this happens, we set these rows to zero.
* To obtain @fig-kernels-gridR2, we make the approximation that the support of $\rho(z_1, z_2)$ is compact in $[-4,4]\times[-4,4]$, which of course is not true, but in the case of a Gaussian distribution this is a reasonable approximation, because $\CDF(4) \approx 0.999968$. Defining $\tilde \rho = \rho \circ S^{-1}$, the grid is numerically renormalized so that despite the compact support approximation, the probabilities sum up to one, and the connectivity matrix is modified so that the columns are weighed by $\tilde \rho(\beta)$:
  $$
  \tilde J_{\alpha\beta} = \tilde \rho(\beta) \sum_{\mu=1}^p \tilde F_{\mu \alpha} \tilde G_{\mu \beta}
  $$
  We see that compared to @fig-kernels-samplingR2, using the grid method allows us to "sample" the locations where the density is small, and this results in less "whited-out" rows. The vertical bands of white correspond to the reweighing of the columns with $\tilde \rho(\beta)$.
* Finally, @fig-kernels-grid01 shows the kernel obtained from applying the grid method in the CDF $[0,1]^2$ described in @sec-mapping-cdf. We see that the connectivity seems much more "uniform", because each point of the grid has "equal weight" in the CDF space.

::: {#fig-kernels layout-ncol=3}

![Sampling neurons $\vec{z_i} \sim \rho(z_1, z_2)$ and binning the corresponding mappings](../chapters/figures/sampling_R2_nogrid_kernel.png){#fig-kernels-samplingR2}

![Grid method in a subset of the PDF space $[-4,4]\times[-4,4] \subset \R^2$](../chapters/figures/sampling_R2_kernel.png){#fig-kernels-gridR2}

![Grid method in the CDF space $[0,1]^2$](../chapters/figures/sampling_R2_01_kernel.png){#fig-kernels-grid01}

(TODO) Comparison of methods for computing the connectivity matrices $\tilde J_{\alpha\beta}$ approximating the kernel $\tilde w(\alpha,\beta)$.
:::

>>>>

::: {.content-hidden}
{{< include ../macros.tex >}}
\def\A{\mathcal{A}}
\def\B{\mathcal{B}}

```{python}
import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.colors as mplcolors, matplotlib.patheffects as pe, matplotlib.ticker as ticker
import numpy as np
import sys
sys.path.append('../../notebooks')  # import neurodyn
from neurodyn import *
from matplotlib_tufte import *
setup()

mpl.use('svg')

def mktrans(fig, ax):
	"""Makes figure and axes transparent"""
	fig.patch.set_facecolor('#FFFFFF00')
	ax.patch.set_facecolor('#FFFFFF00')
```
:::

@@

* @fig-kernels-1 is the method described in this appendix, based on sampling $N$ neurons in the PDF space $\R^2$, applying the mapping on the obtained positions $\vec{z_i}$, the applying the binning (downsampling). We see, through the empty rows in the connectivity, that due to finite-$N$ effect, locations where $\rho$ is small are not represented, and, as a consequence, no neurons are present in the corresponding 1D bin. When this happens, we set these rows to zero.
* To obtain @fig-kernels-2, we make the approximation that the support of $\rho(z_1, z_2)$ is compact in $[-4,4]\times[-4,4]$, which of course is not true, but in the case of a Gaussian distribution this is a reasonable approximation, because $\CDF(4) \approx 0.999968$. Defining $\tilde \rho = \rho \circ S^{-1}$, the grid is numerically renormalized so that despite the compact support approximation, the probabilities sum up to one, and the connectivity matrix is modified so that the columns are weighed by $\tilde \rho(\beta)$:
  $$
  \tilde J_{\alpha\beta} = \tilde \rho(\beta) \sum_{\mu=1}^p \tilde F_{\mu \alpha} \tilde G_{\mu \beta}
  $$
  We see that compared to @fig-kernels-1, using the grid method allows us to "sample" the locations where the density is small, and this results in less "whited-out" rows. The vertical bands of white correspond to the reweighing of the columns with $\tilde \rho(\beta)$.
* Finally, @fig-kernels-3 shows the kernel obtained from applying the grid method in the CDF $[0,1]^2$ described in @sec-mapping-cdf. We see that the connectivity seems much more "uniform", because each point of the grid has "equal weight" in the CDF space.

```{python}
#| label: fig-kernels
#| fig-cap: Comparison of methods for computing the connectivity matrices $\tilde J_{\alpha\beta}$ approximating the kernel $\tilde w(\alpha,\beta)$.
#| fig-subcap: 
#|   - Sampling neurons $\vec{z_i} \sim \rho(z_1, z_2)$ and binning the corresponding mappings.
#|   - Grid method in a subset of the PDF space $[-4,4]\times[-4,4] \subset \R^2$.
#|   - Grid method in the CDF space $[0,1]^2$.
#| layout-ncol: 3
#| echo: false

import matplotlib.colors as mplcolors

nrec = 5

def plot_Jab(Jab):
	amp = np.abs(J_ab).max()
	fig, ax = plt.subplots(dpi=200)
	im = ax.imshow(J_ab,
		cmap='seismic', norm=mplcolors.TwoSlopeNorm(vcenter=0.0, vmin=-amp, vmax=amp),
		extent=((0,1,0,1)), rasterized=True, origin='upper'
		)
	ax.set(
		xlim=(0,1), ylim=(0,1), aspect='equal'
	)
	fig.colorbar(im)
	mktrans(fig, ax)
	breathe(ax)
	return fig, ax

J_ab = LowRankRNNParams.new_valentin(
	p=2, N=100_000, phi=sigmoid, I_ext=zero, exclude_self_connections=False
).to_binmapped(ZMapping.new_nrec(nrec=nrec)).to_dense().J
fig, ax = plot_Jab(J_ab)
display(fig)

phi = sigmoid
rho = lambda z: 1/np.sqrt(2*np.pi)*np.exp(-np.linalg.norm(z, axis=1)**2/2)
z0 = np.random.default_rng(42).normal(loc=0, scale=1, size=1_000_000)
phi_z0 = phi(z0)
a, c = np.mean(phi_z0), np.var(phi_z0)
def phi_tilde(z):
	return (phi(z) - a) / c
mapping = ZMapping.new_nrec(nrec=nrec)  # 4**n = 4**4 = 256 boxes along each axis
Z = mapping.inverse_samples(bbox=Box(xmin=-4, xmax=4, ymin=-4, ymax=4), centered=True)
J_ab = LowRankRNNParams(
	phi=sigmoid, I_ext=zero, exclude_self_connections=False,
	F=Z, G=phi_tilde(Z) * rho(Z)[:, None]/rho(Z).sum() * len(Z)
).to_dense().J
fig, ax = plot_Jab(J_ab)
display(fig)

mapping = ZMapping.new_nrec(nrec=nrec)
V = mapping.inverse_samples()
J_ab = LowRankRNNParams.new_sampled_valentin(
	phi=sigmoid, mapping_samples=V, exclude_self_connections=False, I_ext=zero
).to_dense().J
fig, ax = plot_Jab(J_ab)
display(fig)
```

### 2023-06-20 manuscript/chapters/2_neural_fields.qmd
so... many.... figures !

	axes[i, 1].set_aspect('equal')
	axes[i, 1].set_xlabel('$z_1$')
	axes[i, 1].set_ylabel('$z_2$')
	im = axes[i, 1].pcolormesh(
		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.I_rec(0, rnn.F[:, 0]).reshape(shape),
		cmap='coolwarm', vmin=-2, vmax=2,
		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
	)
	breathe(axes[i, 1])
	if i == 1:
		plt.colorbar(im, ax=axes[i, 1], orientation='horizontal', label='$I^\mathrm{rec}(z_1, z_2)$')

	axes[i, 2].set_aspect('equal')
	axes[i, 2].set_xlabel('$z_1$')
	axes[i, 2].set_ylabel('$z_2$')
	im = axes[i, 2].pcolormesh(
		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.dh(0, rnn.F[:, 0]).reshape(shape),
		cmap=cmap_pan, vmin=-2, vmax=2, shading='gouraud',
		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
	)
	breathe(axes[i, 2])
	if i == 1:
		plt.colorbar(im, ax=axes[i, 2], orientation='horizontal', label='$\partial_t h(z_1, z_2)$')

>>>>

	axes[i, 1].set_aspect('equal')
	axes[i, 1].set_xlabel('$z_1$')
	axes[i, 1].set_ylabel('$z_2$')
	im = axes[i, 1].pcolormesh(
		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.I_rec(0, rnn.F[:, 0]).reshape(shape),
		cmap=cmap_pan, vmin=-2, vmax=2,
		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
	)
	breathe(axes[i, 1])
	if i == 1:
		plt.colorbar(im, ax=axes[i, 1], orientation='horizontal', label='$I^\mathrm{rec}(z_1, z_2)$')

	axes[i, 2].set_aspect('equal')
	axes[i, 2].set_xlabel('$z_1$')
	axes[i, 2].set_ylabel('$z_2$')
	im = axes[i, 2].pcolormesh(
		rnn.F[:, 0].reshape(shape), rnn.F[:, 1].reshape(shape), rnn.dh(0, rnn.F[:, 0]).reshape(shape),
		cmap='coolwarm', vmin=-2, vmax=2, shading='gouraud',
		rasterized=True  # enable rasterization to prevent 50Mo SVG files, lol
	)
	breathe(axes[i, 2])
	if i == 1:
		plt.colorbar(im, ax=axes[i, 2], orientation='horizontal', label='$\partial_t h(z_1, z_2)$')

### 2023-06-20 manuscript/chapters/3_mappings.qmd
so... many.... figures !

mpl.use('svg')
```
:::

@@

::: {#fig-averaging-geom layout-ncol=2}

![Z-mapping](figures/averaging_z.png)

![Column mapping](figures/averaging_column.png)

(TODO) The corresponding square populations corresponding to the segment populations before coarse-graining.
:::

@@

1. The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality. The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out. This results in the erasure of the cycling behavior in the latent trajectory.
2. The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding. Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
3. The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents. In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field.

@@

![(TODO animation, higher N, remove all the other mappings) Simulations of coarse-graining with different mappings from the $p=2$ embedding.](figures/coarsegraining.png){width=500px #fig-cg-2d-1d}

@@

We close this chapter by repeating the process to go from $[0,1]^2$ to $[0,1]$, mapping $(v_1, \alpha_{23})$ to $\alpha_{123}$, and coarse-graining. In total, we have coarse-grained two times: once from $[0,1]^3$ ($8^n$ cube populations) to $[0,1]^2$ ($4^n$ square populations), and again from $[0,1]^2$ to $[0,1]$ ($2^n$ segment populations). The corresponding 1D neural field is now fully fractal, and again exhibits the same dynamics. This shows that "local mappings conserve regularity": we started with a fully continous neural field in $[0,1]^3$, and obtained a partially fractal neural field in $[0,1]^2$. Despite the fractal structure, the 2D neural field is still regular enough that the dynamics are conserved, but more importantly that it can *again* be mapped to an equivalent 1D neural field. In @fig-averaging-geom-3d, we repeat a similar geometric view as @fig-averaging-geom: we see the Z-mapping in the slice $(v_2, v_3)$, how after coarse-graining the kernel is still effectively "sampled" at averaged locations, and crucially, how repeating the coarse-graining still leads to average locations that span the entire $[0,1]^3$ box (instead of being constrained to a plane embedded in $[0,1]^3$, as would have been the case with the Column mapping).

![(TODO what $n$) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px #fig-averaging-geom-3d}

>>>>

mpl.use('svg')

def mktrans(fig, ax):
	"""Makes figure and axes transparent"""
	fig.patch.set_facecolor('#FFFFFF00')
	ax.patch.set_facecolor('#FFFFFF00')
```
:::

@@

```{python}
#| label: fig-averaging-geom
#| fig-cap: Square populations corresponding to the segment populations before coarse-graining are colored identically when they are the same bin of 1D length $2^{-n}$. Here we show $n=3$.
#| fig-subcap:
#|   - Z-mapping
#|   - Column mapping
#| layout-ncol: 2
#| echo: false

from itertools import cycle

def plot_averaging_geom(mapping_cls, nrec=3):
	mapping = mapping_cls.new_nrec(nrec)
	F = mapping.inverse_samples(centered=True)
	F_ = F.reshape((2**nrec, 2**nrec, 2))  # bins before coarse-graining
	values = mapping(F)

	fig, ax = plt.subplots()
	ax.plot(*F.T, color='gray')
	colors = cycle(['#750087','#001997','#06bd01','#ffff00','#fe941e','#ff1e26','#fcaec5','#74d4ea','#603814','#010101','#fbf7f4'])
	for i in range(2**nrec): ax.scatter(*F_[i, :, :].T, zorder=10, color=next(colors), ec='k')
	ax.set(
		xlim=(0,1), ylim=(0,1),
		xlabel='$v_1$', ylabel='$v_2$', aspect='equal',
		xticks=[0, 0.25, 0.5, 0.75, 1], yticks=[0, 0.25, 0.5, 0.75, 1])
	mktrans(fig, ax)
	breathe(ax)
	return fig, ax

fig, ax = plot_averaging_geom(ZMapping, nrec=3)
display(fig)

fig, ax = plot_averaging_geom(ColumnMapping, nrec=3)
display(fig)
```

@@

* @fig-cg-2d-1d-random: The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality. The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out. This results in the erasure of the cycling behavior in the latent trajectory.
* @fig-cg-2d-1d-column: The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding. Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
* @fig-cg-2d-1d-z: The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents. In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field in @fig-cg-2d-1d-original.

@@

::: {#fig-cg-2d-1d}

![Neural field in the 2D embedding](figures/coarse_graining_p=2_original.mp4){#fig-cg-2d-1d-original}

![Random mapping followed by coarse-graining](figures/coarse_graining_p=2_random.mp4){#fig-cg-2d-1d-random}

![Column mapping followed by coarse-graining](figures/coarse_graining_p=2_column.mp4){#fig-cg-2d-1d-column}

![Z-mapping followed by coarse-graining](figures/coarse_graining_p=2_z.mp4){#fig-cg-2d-1d-z}

The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
:::

@@

We close this chapter by repeating the process to go from $[0,1]^2$ to $[0,1]$, mapping $(v_1, \alpha_{23})$ to $\alpha_{123}$, and coarse-graining. In total, we have coarse-grained two times: once from $[0,1]^3$ ($8^n$ cube populations) to $[0,1]^2$ ($4^n$ square populations), and again from $[0,1]^2$ to $[0,1]$ ($2^n$ segment populations). The corresponding 1D neural field is now fully fractal, and again exhibits the same dynamics. This shows that "local mappings conserve regularity": we started with a fully continous neural field in $[0,1]^3$, and obtained a partially fractal neural field in $[0,1]^2$. Despite the fractal structure, the 2D neural field is still regular enough that the dynamics are conserved, but more importantly that it can *again* be mapped to an equivalent 1D neural field.
<!-- In @fig-averaging-geom-3d, we repeat a similar geometric view as @fig-averaging-geom: we see the Z-mapping in the slice $(v_2, v_3)$, how after coarse-graining the kernel is still effectively "sampled" at averaged locations, and crucially, how repeating the coarse-graining still leads to average locations that span the entire $[0,1]^3$ box (instead of being constrained to a plane embedded in $[0,1]^3$, as would have been the case with the Column mapping). -->

<!-- ![(TODO what $n$) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px #fig-averaging-geom-3d} -->

### 2023-06-20 manuscript/chapters/4_locality.qmd
so... many.... figures !

@@

![(TODO change title, change name of mappings, remove the theory, but keep upper bound) (Numerator of the) average binned variation as a function of $n$. The local mappings have a growth slower than the upper bound $2\times 2^n$.](figures/v_num.png){width=500px #fig-vnum}

>>>>

```{python}
#| label: fig-distance-2d-1d
#| fig-cap: Distance between the first population and all other populations, in 1D and 2D. The dotted line has a slope $\sqrt 2$.
#| echo: false

mapping_clss = {
	'Random mapping': RandomMapping,
	'Column mapping': ColumnMapping,
	'Z-mapping': ZMapping,
}

fig, axes = plt.subplots(ncols=len(mapping_clss), figsize=(len(mapping_clss)*3, 3))

isfirst = True
for ax, (name, mapping_cls) in zip(axes, mapping_clss.items()):
	F = mapping_cls.new_nrec(6).inverse_samples()
	a = np.arange(len(F)) / len(F)
	a_ = np.linspace(-0.02, 1.02, 3)
	ax.plot(a, np.linalg.norm(F[0] - F, axis=1), linewidth=0.5)
	ax.plot(a_, a_*np.sqrt(2), '--', clip_on=False) 
	ax.set_xlabel('Distance in 1D')
	if isfirst:
		isfirst = False
		ax.set_ylabel('Distance in 2D')
	else:
		ax.yaxis.set_visible(False)
		ax.spines.left.set_visible(False)
	ax.label_outer()
	ax.set_title(name)
	ax.set_xlim((0,1))
	ax.set_ylim((0, 1.6))
	mktrans(fig, ax)
	breathe(ax)

# %matplotlib inline
display(fig)
```

@@

```{python}
#| label: fig-vnum
#| fig-cap: (Numerator of the) average binned variation as a function of $n$. The local mappings have a growth slower than the upper bound $2\times 2^n$.
#| echo: false

def V(F: np.ndarray) -> np.ndarray:
	nrec = int(np.emath.logn(4, len(F)))
	F_ = F.reshape((2**nrec, 2**nrec, 2))
	# F_ = np.roll(F, 2**nrec).reshape((2**nrec, 2**nrec, 2))  # offset the bins halfway to test
	return (F_.max(axis=1) - F_.min(axis=1)).sum(axis=0)

mapping_clss = {'Random': RandomMapping, 'Column': ColumnMapping, 'Z': ZMapping}
mapping_Vs = { key: [] for key in mapping_clss }
nrecs = np.array([1,2,3,4,5,6,7,8,9])

for name, mapping_cls in mapping_clss.items():
	for nrec in nrecs:
		F = mapping_cls.new_nrec(nrec).inverse_samples(centered=False)
		mapping_Vs[name].append(V(F))

import itertools
fig, ax = plt.subplots(figsize=(4, 3))
marker = itertools.cycle(('s', 'd', '^', 'o', 'P'))
colors = itertools.cycle(('C2', 'C1', 'C0'))
ax.plot(nrecs, 2**(nrecs+1), color='k', linestyle='--', label='upper bound', clip_on=False, zorder=10)
for name, Vs in mapping_Vs.items():
	ax.plot(nrecs, np.array(Vs).sum(axis=-1), marker=next(marker), color=next(colors), label=name, clip_on=False)
ax.set_xlabel('$n$')
ax.set_yscale('log', base=2)
ax.set_autoscaley_on(False)
# ax.grid(True, clip_on=False)
# ax.legend(loc='upper left', bbox_to_anchor=(1,1))
ax.legend(loc='upper left', borderpad=0, borderaxespad=0)
ax.set_xlim((1, 9))
ax.set_xticks([1,3,5,7,9])
ax.set_ylim((1, 2**12))
breathe(ax, which='y')
ax.spines.bottom.set_position(('data', 0.5))  # breathe does not work on log plots
mktrans(fig, ax)
display(fig)
```
